Towards a standardized evaluation of imputation methodology

### 2022-11-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
submitted revisions round 2

\begin{document}
%\DOIsuffix{bimj.DOIsuffix}
\DOIsuffix{bimj.200100000}
\Volume{52}
\Issue{61}
\Year{2010}
\pagespan{1}{}
\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\ [1pc]
% \noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
% } %%% semicolon and fullpoint added here for keyword style
\noindent \hspace*{-4pc} Supporting Information for this article is available from \underline{www.gerkovink.com/evaluation}.\\
% \hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
}  %%% semicolon and fullpoint added here for keyword style

@@

We highly value the additional flexibility that the above outlined design offers. Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against. Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges. On the other hand, we want to warn simulators against potential mishaps, e.g. when calculating and comparing standard errors.

@@

Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's (1976) \nocite{rubi76} definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known. Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not universally sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies, while it offers an informative reference condition in many simulation setups.

@@

At minimum, imputation method performance should be quantified using appropriate measures. It depends on the specifics of each simulation study which performance measures would be most suitable. If the goal is inference--not prediction--the uncertainty about estimates needs to be properly quantified. To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar(Q)$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied. We therefore recommend evaluating the following points:

@@

(iv) Resemblence to the true data values may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may inflate the type I error rate of statistical inferences \citep[Chapter 2.6][]{buur18}. However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between methods. Then, the RMSE may yield valuable information about the methods' comparative efficiency in terms of accuracy and precision.

>>>>

\begin{document}
%\DOIsuffix{bimj.DOIsuffix}
\DOIsuffix{bimj.200100000}
\Volume{52}
\Issue{61}
\Year{2010}
\pagespan{1}{}
\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\ [1pc]
% \noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
% } %%% semicolon and fullpoint added here for keyword style
\noindent \hspace*{-4pc} Supporting information and a simulation archive are available from \underline{www.gerkovink.com/evaluation}.\\
% \hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
}  %%% semicolon and fullpoint added here for keyword style

@@

We highly value the additional flexibility that the above-outlined design offers. Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against. Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges. On the other hand, we want to warn simulators against potential mishaps, e.g. when calculating and comparing standard errors.

@@

Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's (1976) \nocite{rubi76} definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known. Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not universally sound. Sadly, the straightforward case of MCAR is often neglected in simulation studies, while it offers an informative reference condition in many simulation setups.

@@

At minimum, imputation method performance should be quantified using appropriate measures. It depends on the specifics of each simulation study which performance measures would be most suitable. If the goal is inference--not prediction--the uncertainty about estimates needs to be properly quantified. To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar{Q}$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied. We therefore recommend evaluating the following points:

@@

(iv) Resemblance to the true data values may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may inflate the type I error rate of statistical inferences \citep[Chapter 2.6][]{buur18}. However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the prediction errors (residuals) can be compared between methods. Then, the RMSE may yield valuable information about the methods' comparative efficiency in terms of accuracy and precision.
